---
title: "OysterCode"
author: "Ruby K"
date: '2022-06-18'
output: html_document
---

Return to <a href="https://rmk118.github.io/oyster-study" target="_blank">Oyster Study Progress</a>

To see all code and the raw output of any significance tests, see the <a href="https://rmk118.github.io/oyster-study/OysterCode.html" target="_blank">OysterCode</a> page


## Contents

- [Initial Shell Dimensions](#initial-shell-dimensions)
- [Outside Floating Bag Replacement](#outside-floating-bag-replacement)
- [First Sampling Day](#first-sampling-day)

## Initial Shell Dimensions

```{r, initialHeights}
initialHeights1<-c(63,32,48,61,71,41,58,69,33,34,70,61,61,51,39,45,61,60,61,67,60,55,40,69,55,29,60,46,46,53,75,42,21,35,72,61,48,46,47,73,55,31,43,37,52,56,52,40,63,48,50,19,59,59,47,63,41,41,26,35,48,64,44,60,40,49,47,35,55,45,60,42,42,46,49,56,48,45,40,37,60,59,37,28,51,61,45,35,26,54,41,44,58,39,41,47,34,47,50,52,55,43,44,45,41,26,50,38,41,58,42,37,36,48,28,24,32,29,45,70,64,67,24,35,63,56,41,37,47,55,38,75,47,43,40,39,65,50,42,42,40,60,28,36,60,64,48,48,50,49,33,34,40,43,59,46,30,38,58,43,45,35,34,37,48,58,58,42,46,57,44,24,40,40,41,37,59,57,68,40,45,49,50,35,35,30,36,30,32,70,34,41,54,29,43,52,32,37,38,70,42,32,56,37,35,58,50,38,42,43,55,60,43,68,46,60,53,55,59,55,32,50,41,64,37,59,57,52,55,50,38,60,61,37,60,62,55,45,63,47,35,62,61,65,36,62,59,42,39,49,35,50,49,46,43,45,34,44,44,30,41,49,45,41,42,48,32,80,69,69,50,46,57,45,59,63,56,51,45,51)
```

```{r}
oysterAvgInitial<-mean(initialHeights1)
oysterAvgInitial
initialHeightSD<-sd(initialHeights1)
initialHeightSD
shapiro.test(initialHeights1)
```

```{r}
upperLimit<-oysterAvgInitial+2*initialHeightSD
upperLimit
lowerLimit<-oysterAvgInitial-2*initialHeightSD
lowerLimit
```

```{r}
initialHeights<-data.frame(initialHeights1)
```

The initial shell heights were normally distributed (Shapiro-Wilk test, W=0.988, P=0.024). The mean was `r round(oysterAvgInitial, digits=2)` mm and the standard deviation was `r round(initialHeightSD, digits=2)` mm. The acceptable range of oyster shell heights for the experimental gear was set at two standard deviations above and below the mean: `r round(upperLimit, digits=2)` mm to `r round(lowerLimit, digits=2)` mm.

```{r warning=FALSE, initialShellHeightDist, fig.width=4, fig.height=3, fig.align='center'}
library(ggplot2)
ggplot(initialHeights,aes(x=initialHeights1))+geom_histogram(binwidth = 5, fill="#69b3a2", color="#e9ecef")+theme_classic()+xlab("Shell height")+ylab("Frequency")
```
Figure #1: Initial distribution of shell heights in population used to stock experimental gear (n=280)

## Outside Floating Bag Replacement

At some point during the initial sampling day (6/14), the middle floating bag outside the pound became detached from the line and could not be recovered. A replacement bag was stocked from the "leftover" oysters in the population used to stock the original experimental gear. These leftover oysters were being stored in a submerged plastic crate attached to the dock within the North Pound. Only 7 full days had elapsed since the initial stocking was completed, so it was hypothesized that the oysters in the source population would not yet show significant morphological differences from those in the experimental gear.

The replacement ADPI bag was filled with 175 oysters selected in the same manner as the initial treatments (i.e., randomly chosen from the larger population, excluding any animals that were dead, strongly fused together, or had a shell height outside the range of `r round(upperLimit, digits=2)`-`r round(lowerLimit, digits=2)` mm).

Before the placement bag was deployed, a random sample of 32 oysters was collected for comparison to the samples (n=32 for each bag) taken the previous day from the three initial outside ADPI bags. Shapiro-Wilk and Levene's tests were employed to assess the ANOVA assumptions of normality and homoscedasticity, respectively. 

No assumptions were violated for height and length, so a one-way ANOVA was used to determine if significant differences existed between any of the four outside ADPI bags— the three original bags plus the replacement bag.

The shell width distributions were non-normal for both the original bag #2 (p=0.0231) and the replacement bag #2 (p=0.04589). As such, the nonparametric Kruskal-Wallis test was used in place of the ANOVA.

Testing height for homoscedasticity (no violation)
```{r message=FALSE, warning=FALSE}
SamplingOne<-read.csv("6_14.csv")
FBout<-SamplingOne[SamplingOne$Treatment == "FBo",]
library(car)
NewBag<-as.character(FBout$Bag)
FBout$Bag<-NewBag
height.levene<-leveneTest(Height~Bag, data=FBout)
```

Testing height for normality (no violations)
```{r}
bag1height<- FBout[FBout$Bag == "1","Height"]
bag1height.normality<-shapiro.test(bag1height) #Normal
bag2height<- FBout[FBout$Bag == "2","Height"]
bag2height.normality<-shapiro.test(bag2height) #Normal
bag3height<- FBout[FBout$Bag == "3","Height"]
bag3height.normality<-shapiro.test(bag3height) #Normal
bag4height<- FBout[FBout$Bag == "Rep","Height"]
bag4height.normality<-shapiro.test(bag4height) #Normal
```

Height ANOVA: no significant differences
```{r}
HeightResults<-aov(Height~Bag, data=FBout)
summary(HeightResults)
```

Testing length for homoscedasticity (no violation)
```{r message=FALSE, warning=FALSE }
length.levene<-leveneTest(Length~Bag, data=FBout)
```

Testing length for normality (no violations)
```{r}
 
bag1length<- FBout[FBout$Bag == "1","Length"]
shapiro.test(bag1length) #Normal
bag2length<- FBout[FBout$Bag == "2","Length"]
shapiro.test(bag2length) #Normal
bag3length<- FBout[FBout$Bag == "3","Length"]
shapiro.test(bag3length) #Normal
bag4length<- FBout[FBout$Bag == "Rep","Length"]
shapiro.test(bag4length) #Normal
```

Length ANOVA: no significant differences
```{r}
LengthResults<-aov(Length~Bag, data=FBout)
summary(LengthResults)
```

Testing width for homoscedasticity (no violation)
```{r message=FALSE, warning=FALSE }
leveneTest(Width~Bag, data=FBout)
```

Testing width for normality (assumption violated)
```{r}
bag1width<- FBout[FBout$Bag == "1","Width"]
shapiro.test(bag1width) #Normal
bag2width<- FBout[FBout$Bag == "2","Width"]
shapiro.test(bag2width) #Not normal, p=0.0231
bag3width<- FBout[FBout$Bag == "3","Width"]
shapiro.test(bag3width) #Normal
bag4width<- FBout[FBout$Bag == "Rep","Width"]
shapiro.test(bag4width) #Not normal, p=0.04589
```

Width Kruskal-Wallis: no significant differences
```{r}
kruskal.test(Width~Bag, data=FBout)
```

A one-way ANOVA is considered robust to violations of the normality assumption (e.g., Schmider et al., 2010). If an ANOVA is run on the width data despite the noted normality violations, again no significant differences are found.

```{r}
WidthResults<-aov(Width~Bag, data=FBout)
summary(WidthResults)
```

Conclusion: no significant differences were found between the height, length, or widths of the oysters in any of the outside ADPI bags, including the replacement bag. For future analysis of growth and shell morphology, the replacement bag will be analyzed as if it were the original outside bag #2. However, biofouling did appear to be more severe on the replacement oysters than on the oysters in the original outside ADPI bags. As such, when oysters are selected from each treatment for biofouling analysis at the conclusion of the experiment, the outside FB sample will be collected from only bags #1 and #3.

### Within-bag Variability

Within-bag or within-cage variability for each floating bag and cage treatment in given location were assessed using a one-way ANOVA or, if assumptions were not met, a Kruskal-Wallis test. Within-bag variability for outside FB is described above.

##### Inside Floating Bags
```{r}
FBin<-SamplingOne[SamplingOne$Treatment == "FBi",]
#Testing variances (all good)
leveneTest(Height~Bag, data=FBin)
leveneTest(Length~Bag, data=FBin)
leveneTest(Width~Bag, data=FBin)

#Testing height for normality (assumption violated)
bag1height_in<- FBin[FBin$Bag == "1","Height"]
shapiro.test(bag1height_in)
bag2height_in<- FBin[FBin$Bag == "2","Height"]
shapiro.test(bag2height_in)
bag3height_in<- FBin[FBin$Bag == "3","Height"]
shapiro.test(bag3height_in) #p=0.04558

#Testing length for normality (all good)
bag1length_in<- FBin[FBin$Bag == "1","Length"]
shapiro.test(bag1length)
bag2length_in<- FBin[FBin$Bag == "2","Length"]
shapiro.test(bag2length_in)
bag3length_in<- FBin[FBin$Bag == "3","Length"]
shapiro.test(bag3length_in)

#Testing width for normality (assumptions violated)
bag1width_in<- FBin[FBin$Bag == "1","Width"]
shapiro.test(bag1width_in) #Bag 1 not normal, p<0.001
bag2width_in<- FBin[FBin$Bag == "2","Width"]
shapiro.test(bag2width_in) #Bag 2 not normal, p=0.02344
bag3width_in<- FBin[FBin$Bag == "3","Width"]
shapiro.test(bag3width_in) #Bag 3 good
```

```{r}
#Testing for significant differences between inside FB
FBin.height.interbag<-kruskal.test(Height~Bag, data=FBin)
FBin.length.interbag<-aov(Length~Bag, data=FBin)
summary(FBin.length.interbag)
FBin.width.interbag<-kruskal.test(Width~Bag, data=FBin)
```

There were no significant differences in shell height between the three inside floating bags (Kruskal-Wallis test, H(2)=`r round(FBin.height.interbag$statistic, digits=3)`, p=`r round(FBin.height.interbag$p.value, digits=3)`). Similarly, no significant differences in shell width were observed between the three bags (Kruskal-Wallis test, H(2)=`r round(FBin.width.interbag$statistic, digits=3)`, p=`r round(FBin.width.interbag$p.value, digits=3)`).

```{r fig.show="hold", out.width="50%"}
ggplot(data = FBin, aes(x = Bag, y = Height))+geom_boxplot()+theme_classic()+xlab("Bag")+ylab("Height (mm)")+theme(text = element_text(size=15))+scale_y_continuous(limits=c(0,75))

ggplot(data = FBin, aes(x = Bag, y = Width))+geom_boxplot()+theme_classic()+xlab("Bag")+ylab("Width (mm)")+theme(text = element_text(size=15))+scale_y_continuous(limits=c(0,25))
```
Figure #3: Shell height and width of oysters in outside floating bags
<br>

However, the three inside floating bags did significantly differ with regards to shell length (One-way ANOVA, F(2,93)=3.607, p=0.031).
```{r fig.width=4, fig.height=3, fig.align='center'}
ggplot(data = FBin, aes(x = Bag, y = Length))+geom_boxplot()+theme_classic()+xlab("Bag")+ylab("Length (mm)")+theme(text = element_text(size=15))+scale_y_continuous(limits=c(0,50))
```
Figure #4: Shell length of oysters in inside floating bags


### Graphical Results

```{r fig.width=4, fig.height=3, fig.align='center'}
SamplingNew<-SamplingOne[c(1:320,353:608),]
ggplot(data = SamplingNew, aes(x = Gear, y = Height, fill=Location))+geom_boxplot()+scale_y_continuous(limits=c(0,75))
```
Figure #5: Oyster shell height across gear types and locations (n=32 oysters per gear/location combination).

```{r fig.width=4, fig.height=3, fig.align='center'}
ggplot(data = SamplingNew, aes(x = Gear, y = Length, fill=Location))+geom_boxplot()+scale_y_continuous(limits=c(0,50))
```
Figure #5: Oyster shell length across gear types and locations (n=32 oysters per gear/location combination).

```{r fig.width=4, fig.height=3, fig.align='center'}
ggplot(data = SamplingNew, aes(x = Gear, y = Width, fill=Location))+geom_boxplot()+scale_y_continuous(limits=c(0,25))
```
Figure #7: Oyster shell width across gear types and locations (n=32 oysters per gear/location combination).
